{\LARGE Basis Choice for Scalar-on-Function Regression \\ with Applications to Near-Infrared Spectroscopy}

### 2022-02-09 Paper/main.tex
Small changes before meeting

	Functional Data Analysis (FDA), which has its roots in the work of Ulf Grenander and Kari Karhunen, is gaining more attention as researchers from different fields collect data that is functional in nature. Although classical statistical methods can often process this data, FDA has advantages in that it allows extracting information given by properties such as the smoothness of the underlying process or its derivatives (cf. \cite{levitin_introduction_2007}).	As \cite{kokoszka_introduction_2017} describe, using methods from FDA should be considered when one variable of a given data set can be seen as smooth curves or functions.	 
	Therefore, data sets in FDA can include both realizations of scalar random variables and realizations of random functions. Examples of such curves are the absorption curves of light in the Near-infrared (NIR) spectrum by chemical samples.\footnote{For more details on Near-Infrared-Spectroscopy, refer to the Section on Near-Infrared-Spectroscopy in the Appendix.}

@@

	\paragraph{B-spline Basis}
	For $f_1,Y_1$, the MSPE suggests better models with a higher number of FPC's. While the number of basis functions stays at five for $n_{FPC} = 2,3$, it increases to 6 basis functions for $n_{FPC} = 4$. In $f_1,Y_2$, the cross-validated MSPE is only slightly affected by $n_{FPC}$, but lowest for $n_{FPC}$ = 3, indicating the possibility of over-fitting using $n_{FPC}$ = 4. Similar to the two setups with $f_1$, the cross-validated number of basis functions for $f_2$ is increasing in the number of FPC's (4 ,6 ,23 for $n_{FPC} \in \{2, 3, 4 \}$)
	\vspace{-0.2cm}

	\paragraph{Fourier Basis}
	In $f_1,Y_1$, the MSPE is decreasing in $n_{FPC}$ while the results in $f_1,Y_2$ might indicate overfitting when using $n_{FPC}$ = 4. $f_2,Y_1$ displays the greatest relative decline of MPSE in the increasing number of FPC's, but in absolute values, therefore acknowledging the higher noise responses for $f_2,Y2$, the decline of MSPE is similar to the decline observed for $f_2,Y_2$. Both configurations of $f_2$ are using 5, 15, 7 basis functions for $n_{FPC} \in \{2, 3, 4 \}$.
	\vspace{-0.2cm}

@@

	This convex behavior was not observed for the FPCR where no clear relationship between the basis functions and the number of FPC could be uncovered. An hypothesis for this is that the dependency between the truncation $L$ and the FPC's, which are calculated from $L$-truncated smoothed curves, is to complex to capture in our setting. To examine this further, additional simulations were conducted for B-spline basis functions with a large truncation parameter $L \in \{50, 70\}$ for $n_{FPC} \in \{2, \dots, 7 \}$ (LINK TO APPENDIX). As in the simulation study we can find potential signs of overfitting for all setups and both choices of $L$. The additional simulations revealed evidence that there exists a relationship from $n_{FPC} = 3$ onwards: For $n_{FPC} \in \{3, \dots, 7 \}$, the 10-fold cross-validated MSPE is always lower than for 70 basis functions which might be a first sign of a too high $K$ and subsequent undersmoothing. Following this, it seems that once a sufficient number of basis functions is used to expand the curves, the FPCR performs better on a lower truncated smoothed curve (here $L$ = 50). Importantly, the settings $f_1,Y_2$ and $f_2,Y_2$ perform worse for these high number of basis functions, while $f_1,Y_1$ and $f_2,Y_1$ show a stronger performance than in the main simulation study, although the signal-to-noise ration of the responses is not affecting the construction of the FPC's which might indicate benefiting attributes of the lower basis functions used in the main simulation.

>>>>

	Functional Data Analysis (FDA), which has its roots in the work of Ulf Grenander and Kari Karhunen, is gaining more attention as researchers from different fields collect data that is functional in nature. Although classical statistical methods can often process this data, FDA has advantages in that it allows extracting information given by properties such as the smoothness of the underlying process or its derivatives (cf. \cite{levitin_introduction_2007}).	As \cite{kokoszka_introduction_2017} describe, using methods from FDA should be considered when one variable of a given data set can be seen as smooth curves or functions.	 
	Therefore, data sets in FDA can include both realizations of scalar random variables and realizations of random functions. Examples of such curves are the absorption curves of electromagnetic radiation in the Near-infrared (NIR) spectrum by chemical samples.\footnote{For more details on Near-Infrared-Spectroscopy, refer to the \hyperref[NIR]{Section on Near-Infrared-Spectroscopy in the Appendix.}}

@@

	\paragraph{B-spline Basis}
	For $f_1,Y_1$, the MSPE suggests better models with a higher number of FPC's. While the number of basis functions stays at five for $n_{FPC} = 2,3$, it increases to 6 basis functions for $n_{FPC} = 4$. In $f_1,Y_2$, the cross-validated MSPE is only slightly affected by $n_{FPC}$, but lowest for $n_{FPC} = 3$, indicating the possibility of over-fitting using $n_{FPC} = 4$. Similar to the two setups with $f_1$, the cross-validated number of basis functions for $f_2$ is increasing in the number of FPC's (4 ,6 ,23 for $n_{FPC} \in \{2, 3, 4 \}$)
	\vspace{-0.2cm}

	\paragraph{Fourier Basis}
	In $f_1,Y_1$, the MSPE is decreasing in $n_{FPC}$ while the results in $f_1,Y_2$ might indicate overfitting when using $n_{FPC} = 4$. $f_2,Y_1$ displays the greatest relative decline of MPSE in the increasing number of FPC's, but in absolute values, therefore acknowledging the higher noise responses for $f_2,Y2$, the decline of MSPE is similar to the decline observed for $f_2,Y_2$. Both configurations of $f_2$ are using 5, 15, 7 basis functions for $n_{FPC} \in \{2, 3, 4 \}$.
	\vspace{-0.2cm}

@@

	This convex behavior was not observed for the FPCR where no clear relationship between the basis functions and the number of FPC could be uncovered. An hypothesis for this is that the dependency between the truncation $L$ and the FPC's, which are calculated from $L$-truncated smoothed curves, is to complex to capture in our setting. To examine this further, additional simulations were conducted for B-spline basis functions with a large truncation parameter $L \in \{50, 70\}$ for $n_{FPC} \in \{2, \dots, 7 \}$ (LINK TO APPENDIX). As in the simulation study we can find potential signs of overfitting for all setups and both choices of $L$. The additional simulations revealed evidence that there exists a relationship from $n_{FPC} = 3$ onwards: For $n_{FPC} \in \{3, \dots, 7 \}$, the 10-fold cross-validated MSPE is always lower than for 70 basis functions which might be a first sign of a too high $K$ and subsequent undersmoothing. Following this, it seems that once a sufficient number of basis functions is used to expand the curves, the FPCR performs better on a lower truncated smoothed curve (here $L = 50$). Importantly, the settings $f_1,Y_2$ and $f_2,Y_2$ perform worse for these high number of basis functions, while $f_1,Y_1$ and $f_2,Y_1$ show a stronger performance than in the main simulation study, although the signal-to-noise ration of the responses is not affecting the construction of the FPC's which might indicate benefiting attributes of the lower basis functions used in the main simulation.
